---
title: "Aides Drug"
author: "Connor Rice"
date: "9/21/2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)

drugusers <- triUmp %>% filter(DrugUser=="yes")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$ There is no difference between the rate of infection between drug users who are given the placebo and those who are given the actual drug.
$H_a$ There is a significant difference in infection rate between drug users who are given the placebo versus those who are given the actual drug.


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Data on the efficacy of the drug in 'drug users' will be done by a two way variable analysis of whether or not they were infected versus whether or not they were on the placebo or the real drug. The full data was initially filtered to only include the data of those who are 'drug users'. The two categorical data types will be analyzed using the cat_cat helper file.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drugusers)
barchartGC(~infected + treatment,data=drugusers, type="percent")

```

The large percentage reduction in the number of infections in the people who are getting the placebo versus those who are getting the actual drug indicates a high potential the drug is effective in preventing HiV/AIDs infection. This needs to be confirmed by further numerical and inferential statisical analysis.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drugusers)
rowPerc(table1)
colPerc(table1)
```

The large reduction in percent in infection rates of 'drug users' between those who received the drug and those who received the placebo(2.89% of those on actual drug were infected versus 7.77% of those who received the placebo) is a good indication that the drug effectively decreases overall infection rates in 'drug users'.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfies the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
fisher.test(table1)
```

The very low p-values in the chi-squared test and fisher exact test(.0000001242 and .0000000423 respectively) indicate that the difference in infection rates is indeed statistically significant which means the drug is effective at reducing infection rates in 'drug users'and allows us to reject the null hypothesis. However the confidence interval of 0.23-0.53 is a bit lower than what would be desired for a preventative drug like a vaccine.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

There is a large significant difference in the data of infection rates of 'drug users' who are on the actual drug versus those who are on the placebo, specifically with a large percent reduction in infection rates in those who are on the actual drug. All forms of data analysis performed support this conclusion and support the rejection of the null hypothesis; however the confidence interval of the drug is somewhat low for what would be ideal in a preventative drug.
